---
title: "Final Project"
author: "Alexis Adams-Clark"
date: "2/13/2019"
output: html_document
---

I really appreciated that you had the code to install less commonly used packages, definitely helps with 
reproducibility! 
```{r setup, include = FALSE}
library(knitr)
library(rio)
library(tidyverse)
library(car)
library(waffle)
library(extrafont)
library(emojifont)


#install.packages("extrafont")
#install.packages("emojifont")
#install.packages("extrafont")
#devtools::install_github("wch/extrafontdb")
#devtools::install_github("wch/Rttf2pt1")
#devtools::install_github("hrbrmstr/waffle")
#font_import()
#loadfonts()
#fonts()[grep("Awesome", fonts())]

#Yeah, I can't get my waffle plot to work with glyphs, so if any of you have recommendations about that, please let me know!!

knitr::opts_chunk$set(fig.width=8, fig.height=5, fig.align = "center", message = FALSE, error = FALSE, warning = FALSE)
theme_set(theme_minimal(base_size = 25))
options(scipen=999)

# I typically have this code chunk separate from library and data import. 

data <- Salaries

data$sex <- factor(data$sex, levels = c("Male", "Female"))

```

```{r plot 1}
#Plot 1 - Goal: Displaying how many professors there are by gender and on average how much they are each paid at University X

data1 <- data %>% 
  group_by(sex) %>% 
  summarize(mean = mean(salary))

#Option 1: Histogram, I like this one because it gives you a sense of numbers (there are not enough female professors), and also a sense of how much less they are paid, but definitely not for a public audience
plot1a <- ggplot(data, aes(salary)) +
  geom_histogram(aes(fill = sex), alpha = 0.7, color= "white") +
  facet_wrap(~sex) + 
  labs(title = "The Gender Gap in Academia:", 
       subtitle = "The Distribution of Faculty Salaries by Gender at University X \n", 
       x = "\nSalary",
       y = "Number of Professors \n") + 
  theme(legend.position="none", 
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5)) + 
  scale_fill_manual(values = c("#66C2A5", "#440154FF")) + 
  scale_x_continuous(labels=scales::dollar)
plot1a


#Option 2: Boxplot, I like that this shows you the mean/af the distributions, but I don't like that you can't see the actual data points and that you can't really get a sense of the discrepancy in the number of female and male professors. Also less understandable by the public.
plot1b <- ggplot(data, aes(x = sex, y = salary, fill = sex)) + 
  geom_boxplot(alpha = 0.7) + 
  labs(title = "The Gender Gap in Academia:", 
       subtitle = "The Distribution of Faculty Salaries by Gender at University X \n",
       y = "Salary\n", 
       x = " ") + 
  theme(legend.position="none",
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5)) + 
  scale_fill_manual(values = c("#66C2A5", "#440154FF")) + 
  scale_y_continuous(labels=scales::dollar)
plot1b


#Final Version/Option 3:
library(scales) #maybe add this up top with other libraries
plot1c <- ggplot(data, aes(x = sex, y = salary, color = sex)) + 
  geom_jitter(width = 0.15, alpha = 0.7, size = 3) + 
  theme(legend.position="none",
         plot.title = element_text(face = "bold", hjust = 0.5),
         plot.subtitle = element_text(face = "bold", hjust = 0.5)) + 
  geom_segment(data = filter(data1, sex == "Male"), aes(x = 0.75, xend = 1.25, y = mean, yend = mean), linetype = "dashed", size = 1, color = "black") + 
  geom_segment(data = filter(data1, sex == "Female"), aes(x = 1.75, xend = 2.25, y = mean, yend = mean), linetype = "dashed", size = 1, color = "black") + 
  annotate("text", label = "Average Salary $115,090", x = 1, y = 225000, color = "#66C2A5", face = "bold", size = 10) +
  annotate("text", label = "Average Salary $101,002", x = 2, y = 225000, color = "#440154FF", face = "bold", size = 10) + 
  labs(y = "Salary \n", x = " ") + 
  scale_color_manual(values = c("#66C2A5", "#440154FF")) + 
  labs(title = "The Gender Gap in Academia:", 
       subtitle = "Faculty Salaries by Gender at University X \n", 
       x = "\n", 
       y = "Salary \n") + # I like this little trick \n to move the labels away from the plot
  scale_y_continuous(labels=scales::dollar)
plot1c

#still need to adjust theme font, think about setting y limit to ylim(0, 240000), and change annotations to look better, add caption with data source
```
Notes for plot1

Code was straight-forward and easy to follow!
I enjoyed watching the transition from plot 1a-1c. You have so many areas of strengths for this plot including: large/easy to read text, helpful labels and title/subtitle. I like your annotations and the use of individual data points because it shows the distribution of data and the lack of female professors. One of my comments would be to extend the y axis which it looks like you are already planning on doing, you could also just have the annotations be black from simplicity! Overall it's great and the message is clear. 
```{r plot 2} 
#Plot 2 Goal: Showing how many female and male professors there at University X, by rank
  
#Option 1: Just a simple stacked bar graph, but as we talked about in class, these kind of stink
plot2a <- ggplot(data, aes(x = sex, fill = rank)) + 
  geom_bar(alpha = 0.8) + 
  scale_fill_viridis_d(name="Rank",
                       labels=c("Assistant Prof", "Associate Prof", "Full Prof")) + 
  labs(title = "University X Doesn't Hire Enough Female Professors", 
       subtitle = "And has a disproportionate number of men in the most powerful positions...\n", 
       y = "Number of Professors \n", 
       x = " ") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5))
plot2a

data2 <- data
data2$sex <- factor(data2$sex, levels = c("Female", "Male"))

#Option 2: Dodged bar graph with coord flip
plot2b <- ggplot(data2, aes(x = sex, fill = rank)) + 
  geom_bar(position = "dodge", alpha = 0.8) + 
  scale_fill_viridis_d(name="Rank",
                       labels=c("Assistant Prof", "Associate Prof", "Full Prof")) + 
  coord_flip() + 
  labs(title = "University X Doesn't Hire Enough Female Professors", 
       subtitle = "And has a disproportionate number of men in the most powerful positions...\n", 
       y = "\n Number of Professors", 
       x = " ") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5))
plot2b

#Final version/Option 3: 2 Waffle Charts
data_summary2 <- data %>% 
  group_by(sex, rank) %>% 
  summarize(n = n())

partsf <- c("Assistant Professor" = 11, "Associate Professor" = 10, "Full Professor" = 18)


plot2cw <- waffle(partsf, 
       rows = 3, 
       title = "Female Professors\n",
       size = 4)
plot2cw + scale_fill_viridis_d()

partsm <- c("Assistant Professor" = 56, "Associate Professor" = 54, "Full Professor" = 247)

plot2cm <- waffle(partsm, 
       rows = 7, 
       title = "Male Professors\n", 
       size = 4)
plot2cm + scale_fill_viridis_d()

# I'm going to figure out a way to merge the two waffles plots and make the squares the same size, but haven't quite figured out how to do that yet
#Also, for the life of me I have not had luck with getting fontAwesome to work on my plot for glyphs, so if anyone has any tips, that would be great!
#Still need to adjust theme font, add caption with data source
```
Notes for plot 2 
As I mentioned for the first plot- watching the transformation is fun! A l0t of strengths for your final plots: using waffle plots to show counts rather than a bar plot, color blind safe palette, clear labels. I only have a few suggestions: change "value" to "Rank" like you had in the first few plots, make legend text a little bigger. You could probably even use annotations to label on the blocks rather than using a legend. I haven't actually needed to merge multiple plots but you could try "par(mfrow=c(1,2))"
```{r plots}

#Plot 3 Goal: Display the discrepancy in salaries between men and women at different professor ranks (assistant professor, associate professor, and full professor)

#Option 1: Bar graph, but apparantly dynamite plots are bad.....! # I honestly don't mind them that much ha!
plot3a <- data %>% 
  group_by(sex, rank) %>% 
  summarize(mean = mean(salary), 
            salary_se = sqrt(var(salary, na.rm = TRUE) / length(salary))) %>% 
  ggplot(aes(x = sex, y = mean, fill = sex)) + 
  geom_col(alpha = 0.8) + 
  geom_errorbar(aes(ymin = (mean - 1.96*salary_se), ymax = (mean + 1.96*salary_se)), 
                color = "gray40",
                width = 0.2,
                size = 0.5) +
  facet_wrap(~rank) + 
  labs(title = "University X Doesn't Pay Female Professors as Much as Male Professors", 
       subtitle = "Mean Salary Faculty Rank...\n", 
       y = "Number of Professors \n", 
       x = " ") + 
  theme(legend.position="none",
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5)) + 
  scale_fill_manual(values = c("#66C2A5", "#440154FF")) + 
  scale_y_continuous(labels=dollar)
plot3a


#Option 2: Similar plot but with points
plot3b <- data %>% 
  group_by(sex, rank) %>% 
  summarize(mean = mean(salary),
            salary_se = sqrt(var(salary, na.rm = TRUE) / length(salary))) %>% 
  ggplot(aes(x = sex, y = mean, color = sex)) + 
  geom_point(size = 5) + 
  geom_errorbar(aes(ymin = (mean - 1.96*salary_se), ymax = (mean + 1.96*salary_se)), 
                color = "gray40",
                width = 0.2,
                size = 0.5) + 
  facet_wrap(~rank) + 
  labs(title = "University X Doesn't Pay Female Professors as Much as Male Professors", 
       subtitle = "Mean Salary Faculty Rank...\n", 
       y = "Number of Professors \n", 
       x = " ") + 
  theme(legend.position="none",
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5)) + 
  scale_color_manual(values = c("#66C2A5", "#440154FF")) + 
  scale_y_continuous(labels=dollar) + 
  ylim(0, 140000)
plot3b

#Option 3: This is alittle bit different take on the same idea, and also adds a different element. It looks at the proportion of men's salaries that the women's salaries represent in the different professor ranks, but ALSO adds another piece of information- the discipline that they aare working in (either theoretical or applied). Thus, highlighting that the problem is really in the theoretical disciplines. 

data <- data %>% 
  mutate(rank = fct_recode(rank, 
                     "Assistant Professor" = "AsstProf",
                     "Associate Professor" = "AssocProf",
                      "Full Professor" = "Prof"),
         discipline = fct_recode(discipline, 
                                 "Theoretical" = "A",
                                 "Applied" = "B"))

plot3c <- data %>% 
  group_by(sex, rank, discipline) %>% 
  summarize(mean = mean(salary)) %>% 
  spread(sex, mean) %>% 
  mutate(prop = (1 - ((Male - Female)/Male)),
         prop = as.numeric(prop)) %>% 
  select(rank, discipline, prop) %>% 
  ggplot(aes(x = rank, y = prop, fill = rank)) + 
  geom_col(alpha = 0.8) + 
  geom_segment(aes(x = 0, xend = 4, y = 1, yend = 1), linetype = "dashed", size = 2, color = "red") + 
  labs(title = "Pay Gap is Significant in Theoretical Departments at University X", 
       subtitle = "Women's Pay Expressed as Percentage of Men's Pay\n", 
       y = "\n% of Men's Pay", 
       x = " ") + 
  theme(legend.position="none",
        plot.title = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(face = "bold", hjust = 0.5)) + 
  scale_fill_viridis_d() + 
  facet_grid(discipline~.) +
  scale_y_continuous(breaks = c(0.25, 0.5, 0.75, 1.0),
                     labels = c("25%", "50%", "75%", "100%"),
                     limits = c(0, 1.25)) +
  coord_flip()
ann_text<-data.frame(rank=c("Assistant Professor","Associate Professor","Full Professor","Assistant Professor","Associate Professor","Full Professor")
                     ,prop=c(1.10,1.10,1.10,1.10,1.10,1.10), discipline=c("Theoretical","Theoretical","Theoretical","Applied","Applied","Applied"),
                     label=c("98.2%","84.8%", "90.8%", "99.4%", "97.8%", "98.7%"))

plot3c + geom_text(data = ann_text,label=ann_text$label, size = 8)

##Still need to adjust theme font, possibly colors, add caption with data source

```
Plot 3 notes:
I think this is a really interesting way to look at this relationship. I think for a scientific audience this is a really good plot but has a lot of parts for a general audience. I'm not quite sure I understand what the little red bars are (is that the line at 100%). The plot does highlight that it seems the issues is in the theoretical discipline! I like that added piece of information; the colors and font size is also great! My only suggestion would be to make % of Men's Pay be "percent of Men's Pay and maybe a solid line instead of the red dotted line, or remove it altogether! 
```{r plot 4}

#Plot 4 Goal: Display the salary of men and women by years of service and years since phd

plot4a <- ggplot(data, aes(x = yrs.service, y = salary, color = sex)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm") + 
  labs(title = "Gender Pay Gap at University X As a Function of Years of Service", 
       subtitle = " ", 
       y = "Salary \n", 
       x = "\nYears of Service") + 
  scale_color_manual(name = "Gender", values = c("#66C2A5", "#440154FF")) + 
  facet_wrap(~discipline) + 
  scale_y_continuous(labels=dollar)
plot4a

plot4b <- ggplot(data, aes(x = yrs.since.phd, y = salary, color = sex)) + 
  geom_point(alpha = 0.6) + 
  geom_smooth(method = "lm") + 
  facet_wrap(~discipline) + 
  labs(title = "Gender Pay Gap at University X As a Function of Years Since Ph.D.", 
       subtitle = " ", 
       y = "Salary \n", 
       x = "\nYears Since Ph.D.") + 
  scale_color_manual(name = "Gender", values = c("#66C2A5", "#440154FF")) + 
  scale_y_continuous(labels=dollar)
plot4b
#Still playing around with, Still need to adjust theme font, add caption with data source

```
Notes for plot4
I think this is a really interesting plot! Once again a lot of strengths including large font size, clear labels, and showing individual data points. My only suggestion would be to maybe consider suppressing the se on the regression line so that you can see individual data points and then it would take away that gray box over the line/points in the legend. 
